\chapter*{Acknowledgement}

This manuscript is an early version of my Ph.D. Thesis. The decision of writing this is not easy. Part of it is due to exhausion about the field of traditional GIS, other parts is the quarantine period caused by the coronavirus in 2020, maybe the \textit{Thesis in the Time of Cholera}. During these months I have been home with my parents, enjoying my mothere's cuisine and patience. I sometimes stay up so late that we can only meet in the late evenings, even we live under the same roof. My father does not really care about what I do and how I manage to do them. I am sorry for not being a good boy who can watch television with my parents everyday but still receive your support. Thank you for being there.

Zheng Yuefang, Zhang Yuqi, Wang Yuhan, and Guo Naiyao have been dear friends of mine since we first got to PKU. Thankfully we still share our hopes and dreams. I love you guys!

Wang Jianying, Xing Xiaoyue, and Gong Xuri, you know about my everything since I first came to IRSGIS. The moments we share is the most precious beyond academics in the \textit{getting-boring} years. 

Zhong Yanjie and Wu Haoyin, it is fancy that we belong to different time zones but seems like none. Maybe because I am too far in night navigation. Unfortunately now I am a Melatonin taker.

The writing of this thesis has been a miracle for me. I cannot imagine how many times I have come up to my computer and start to say to myself, `Hello darkness my old friend, I come to talk to you again', but finally here it comes. I love the peace that doing research has brought me. It is time to take a pause and move on.